---
title: 'How data drives early word learning: Comprehension'
author: "F.Mollica S.T.Piantadosi"
date: "12/09/2016"
output: html_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(warning=FALSE)
suppressMessages(library(ggplot2))
suppressMessages(library(plyr))
suppressMessages(library(reshape))
```

# Comprehension All Data

```{r}
# Load in the data
data = read.csv('All_Data_Comprehension_Summary.csv')
mcdi = read.csv('comprehension-mcdi.csv')
converge = read.csv('All_Data_Comprehension_Convergence.csv')
converge$converged = "No"
converge$converged[converge$Rhat_k < 1.2 & converge$Rhat_l < 1.2 & converge$Rhat_s < 1.2] = 'Yes'

message(paste('Percent Chains Converged:', 100*dim(converge[converge$converged=='Yes',])[1] / dim(converge)[1]))

d = merge(data, converge[c('language','item.id','converged')])
d = subset(d, converged=='Yes')

```


## Check Model Fit

```{r}
pred = merge(mcdi, d[,c('language', 'item.id', 'k_Mean', 'lambda_Mean', 'start_Mean')])

PRED = NULL
for(q in split(pred, paste(pred$language, pred$item))) {
    train <- q
    test  <- q
    train$successes = train$produces * train$n
    train$failures = train$n - train$successes
    
    g.Probit <- glm( cbind(successes, failures) ~ age, family=binomial(link="probit"), data=train)
    g.Logit  <- glm( cbind(successes, failures) ~ age, family=binomial(link="logit"),  data=train)
            
    PRED <- rbind(PRED, data.frame(language=q$language[1], item.id=q$item.id[1], item=q$item[1], category=q$category[1], 
                                   k=q$k_Mean[1], l=q$lambda_Mean[1], s=q$start_Mean[1], obs=test$produces, n=q$n[1], age=test$age,
                                    Gamma=pgamma(test$age-q$start_Mean[1], shape=q$k_Mean[1], rate=q$lambda_Mean[1]),
                                    Probit=predict(g.Probit, newdata=test, type="response"),
                                    Logit=predict(g.Logit, newdata=test, type="response")
                                    ))
}

pred.m <- melt(PRED, id.vars=c('language', 'item.id', 'age', 'obs', 'n', 'category', 'item', 'k', 'l', 's'))

mycor.test <- function(df) {
           cr <- cor.test(df$value, df$obs) 
           data.frame(estimate=cr$estimate, ymin=cr$conf.int[1], ymax=cr$conf.int[2])
 }

m <- ddply( pred.m, c("language", "variable"), mycor.test)

m$Model <- m$variable

ggplot(m, aes(y=estimate, x=language, color=Model, ymin=ymin, ymax=ymax)) + 
        geom_point() + geom_errorbar() + 
        ylim(0,1) + theme_bw() +
        theme(axis.text.x = element_text(angle = 90)) + 
        xlab("") + ylab("Coef. of Determination")
```

## Check Parameter Values

```{r}
mean = ddply(d, .(language), summarise, k_M=mean(k_Mean), k_low=mean(k_HDIlow), k_high=mean(k_HDIhigh), 
                                        l_M=mean(lambda_Mean), l_low=mean(lambda_HDIlow), l_high=mean(lambda_HDIhigh), 
                                        s_M=mean(start_Mean), s_low=mean(start_HDIlow), s_high=mean(start_HDIhigh))
mean
```

### Number of ELIs: k

```{r}
ggplot(d, aes(x=language, y=k_Mean, fill=language)) + 
    geom_boxplot() + 
    ylab("Estimated k") + xlab("") + ylim(0, 65) +
	guides(fill=F) + ggtitle('Comprehension') +
	theme_bw() +
	theme(axis.text.x=element_text(angle=90))
```

### Rate of ELIs: $\lambda$

```{r}
ggplot(d, aes(x=language, y=lambda_Mean, fill=language)) + 
    geom_boxplot() + 
	ylab(expression(paste("Estimated ", lambda))) + xlab("") +
	guides(fill=F) + ggtitle('Comprehension') + ylim(0, 3) +
	theme_bw() +
	theme(axis.text.x=element_text(angle=90))
```

### Start Time: s

```{r}
ggplot(d, aes(x=language, y=start_Mean, fill=language)) + 
    geom_boxplot() + 
	ylab("Estimated start (months)") + xlab("") +
	guides(fill=F) + ggtitle('Comprehension') +
	theme_bw() + ylim(0,20) +
	theme(axis.text.x=element_text(angle=90))
```

## Parameter Values by MCDI Category for English

```{r}
eng = subset(d, language=='English')
```

### Number of ELIs: k

```{r}
ggplot(eng, aes(x=category, y=k_Mean, fill=category)) + 
    geom_boxplot() + 
	ylab("Estimated mean k") + xlab("") +
	guides(fill=F) + ggtitle('Comprehension') +
	theme_bw() +
	theme(axis.text.x=element_text(angle=90))
```

### Rate of ELIs: $\lambda$

```{r}
ggplot(eng, aes(x=category, y=lambda_Mean, fill=category)) + 
    geom_boxplot() + 
	ylab(expression(paste("Estimated mean ",lambda))) + xlab("") +
	guides(fill=F) + ggtitle('Comprehension') +
	theme_bw() +
	theme(axis.text.x=element_text(angle=90))
```

### Start Time: s

```{r}
ggplot(eng, aes(x=category, y=start_Mean, fill=category)) + 
    geom_boxplot() + 
	ylab("Estimated mean start time (months)") + xlab("") +
	guides(fill=F) + ggtitle('Comprehension') +
	theme_bw() +
	theme(axis.text.x=element_text(angle=90))
```

## Correlating Parameter Values with Childes Frequency

```{r}
childes <- read.table("childes.txt", header=F)
names(childes) <- c("item", "freq")
freq <- merge(eng, childes)
```

### Number of ELIs: k

```{r}
r = with(freq, cor.test(log(freq), k_Mean))
message(paste('Correlation with Log Frequency: ', r$estimate, ', p=', r$p.value, sep=''))

ggplot(freq, aes(x=log(freq), y=k_Mean)) + 
    geom_point(alpha=0.5) + 
    stat_smooth(method='lm',formula=y~x) + 
    xlab("Log Frequency") + 
    ylab(expression(paste("Estimated ", k))) + 
    theme_bw()
```

### Rate of ELIs: $\lambda$

```{r}
r = with(freq, cor.test(log(freq), lambda_Mean))
message(paste('Correlation with Log Frequency: ', r$estimate, ', p=', r$p.value, sep=''))

ggplot(freq, aes(x=log(freq), y=lambda_Mean)) + 
    geom_point(alpha=0.5) + 
    stat_smooth(method='lm',formula=y~x) + 
    xlab("Log Frequency") + 
    ylab(expression(paste("Estimated ", lambda))) + 
    theme_bw()
```

### Start Time: s

```{r}
r = with(freq, cor.test(log(freq), start_Mean))
message(paste('Correlation with Log Frequency: ', r$estimate, ', p=', r$p.value, sep=''))

ggplot(freq, aes(x=log(freq), y=start_Mean)) + 
    geom_point(alpha=0.5) + 
    stat_smooth(method='lm',formula=y~x) + 
    xlab("Log Frequency") + 
    ylab(expression(paste("Estimated ", 'start time (months)'))) + 
    theme_bw()
```

## Comparison of Variance Explained by Waiting for Data vs. Start Time

```{r}
bootByHand = function(df, iters) {
    RESULTS = NULL
    for (i in 1:iters) {
        N = length(df$language)
        inx = sample(1:N, N, replace=T)
        set = df[inx,]
        
        R.t = var(set$k_Mean / set$lambda_Mean)
        R.a = var(set$start_Mean) + var(set$k_Mean / set$lambda_Mean)
        PropVar = R.t / R.a

        set$t = set$k_Mean / set$lambda_Mean
        set$a = set$start_Mean + set$t
        PropTime = set$t / set$a
        
        result = data.frame(PropVar = PropVar, PropTime = PropTime)
        RESULTS = rbind(RESULTS, result)
    }
    return(RESULTS)
}

propBootC = ddply(d, .(language), bootByHand, 1000)
prC = ddply(propBootC, .(language), summarise, lowVar = quantile(PropVar, probs=0.05), lowT = quantile(PropTime, probs=0.05), 
            highVar = quantile(PropVar, probs=0.95), highT = quantile(PropTime, probs=0.95))

get_prop = function(d) {
    R.s.comp = var(d$start_Mean)
    R.t.comp = var(d$k_Mean / d$lambda_Mean)
    R.a.comp = var(d$start_Mean) + var(d$k_Mean / d$lambda_Mean)
    Vars=R.t.comp/R.a.comp
    
    set = d
    set$t = set$k_Mean / set$lambda_Mean
    set$a = set$start_Mean + set$t
    propTime = mean(set$t / set$a)
    
    return(data.frame(PropV=Vars, PropT=propTime))
}

proC = ddply(d, .(language), get_prop)
propC = merge(prC, proC)

ggplot(propC, aes(x=language, y=PropV, fill=language)) +
    geom_bar(stat='identity', position=position_dodge(0.9)) +
    geom_errorbar(aes(ymin=lowVar,ymax=highVar), position=position_dodge(0.9), width=0.3) +
    geom_point(aes(x=language, y=PropT), shape=2) +
    geom_pointrange(aes(ymin=lowT, ymax=highT)) +
    ylab('Percent') +
    xlab('') +
    guides(fill=F) +
    ggtitle('Comprehension') +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90))
```

# Comprehension First Half Data

```{r}
# Load in the data
data = read.csv('First_Half_Comprehension_Summary.csv')
converge = read.csv('First_Half_Comprehension_Convergence.csv')
converge$converged = "No"
converge$converged[converge$Rhat_k < 1.2 & converge$Rhat_l < 1.2 & converge$Rhat_s < 1.2] = 'Yes'

message(paste('Percent Chains Converged:', 100*dim(converge[converge$converged=='Yes',])[1] / dim(converge)[1]))

d = merge(data, converge[c('language','item.id','converged')])
d = subset(d, converged=='Yes')

```

## Check Predictive Ability

```{r}
pred = merge(mcdi, d[,c('language', 'item.id', 'k_Mean', 'lambda_Mean', 'start_Mean')])

PRED = NULL
for(q in split(pred, paste(pred$language, pred$item))) {
    train <- subset(q, age<median(age))
    test  <- subset(q, age>=median(age))
    train$successes = train$produces * train$n
    train$failures = train$n - train$successes
    
    g.Probit <- glm( cbind(successes, failures) ~ age, family=binomial(link="probit"), data=train)
    g.Logit  <- glm( cbind(successes, failures) ~ age, family=binomial(link="logit"),  data=train)
            
    PRED <- rbind(PRED, data.frame(language=q$language[1], item.id=q$item.id[1], item=q$item[1], category=q$category[1], 
                                   k=q$k_Mean[1], l=q$lambda_Mean[1], s=q$start_Mean[1], obs=test$produces, n=q$n[1], age=test$age,
                                    Gamma=pgamma(test$age-q$start_Mean[1], shape=q$k_Mean[1], rate=q$lambda_Mean[1]),
                                    Probit=predict(g.Probit, newdata=test, type="response"),
                                    Logit=predict(g.Logit, newdata=test, type="response")
                                    ))
}

pred.m <- melt(PRED, id.vars=c('language', 'item.id', 'age', 'obs', 'n', 'category', 'item', 'k', 'l', 's'))

mycor.test <- function(df) {
           cr <- cor.test(df$value, df$obs) 
           data.frame(estimate=cr$estimate, ymin=cr$conf.int[1], ymax=cr$conf.int[2])
 }

m <- ddply( pred.m, c("language", "variable"), mycor.test)

m$Model <- m$variable

ggplot(m, aes(y=estimate, x=language, color=Model, ymin=ymin, ymax=ymax)) + 
        geom_point(position=position_dodge(.2)) + 
        geom_errorbar(position=position_dodge(.2),width=2) + 
        ylim(0,1) + theme_bw() +
        theme(axis.text.x = element_text(angle = 90)) + 
        xlab("") + ylab("Coef. of Determination") +
        ggtitle('Comprehension')
```
